---
title: "trade_paper"
output: html_document
---
# Turkey-Venezuela Free Trade Agreement and Its Analysis
## Sarp Dora Cebeci - 070170403 (ITU ID)
### sarpdoracebeci@gmail.com

# Session
```{r}
setwd("~/Downloads/Trade_Paper")
library(tidyverse)
library(WDI)
library(cepiigeodist)
library(xlsx)
library(tradestatistics)
library(plm)
library(stargazer)
```

# Datasets
```{r}
# downloaded from: https://www.rug.nl/ggdc/historicaldevelopment/maddison/releases/maddison-project-database-2020
mpd_1 <- read.csv("mmpd.csv")
mpd_1 <- mpd_1 %>% mutate(countrycode = tolower(countrycode))
# from wits.worldbank.org api
turkey_trade <- read_csv("turkey_trade.csv")
# creating gdp and specifying the time period
mpd_2 <- mpd_1 %>% mutate(gdp = gdppc * pop) %>% 
  filter(year > 1961) %>% select(c(year, countrycode, country, gdp))
# creating the total trade and renaming
turkey_trade1 <- turkey_trade %>% 
  mutate(total_trade = (export_value_usd + import_value_usd)/2) %>% 
  select(c(year, partner_iso, partner_fullname_english, total_trade)) %>% 
  rename(code = partner_iso, country = partner_fullname_english)
# gathering the two
df <- left_join(turkey_trade1, mpd_2, by = c("code" = "countrycode",
                                       "year" = "year"))
```

```{r}
# getting turkey's gdp 
tr_gdp <- mpd_1 %>% mutate(gdp = gdppc * pop) %>% 
  filter(country == "Turkey",
         year > 1961) %>% 
  select(gdp)
# and the world's gdp on overall
gdp_world <- WDI(country = "all", indicator = "NY.GDP.MKTP.CD", 
    start = 1962, end = 2018) %>% 
  filter(country == "World") %>%
  select(c("NY.GDP.MKTP.CD","year")) %>%
  rename(w_gdp = NY.GDP.MKTP.CD) %>% 
  arrange(year)
# combining the two
tr_w <- gdp_world
tr_w$tr_gdp <- tr_gdp$gdp
# adding these two
df2 <- left_join(df, tr_w, by = c("year" = "year"))
```

```{r}
# distances from turkey
## distw used which puts weight on the parts of the countries
## economical centers trade centers
dist_tr <- dist_cepii %>% 
  mutate(iso_o = tolower(iso_o),
         iso_d = tolower(iso_d)) %>% 
  filter(iso_o == "tur") %>% 
  select(iso_d, distw)
# adding distance
df3 <- left_join(df2, dist_tr, by = c("code" = "iso_d"))
# dataset with wanted columns 
df4 <- df3 %>% 
  select(c(code, year, total_trade, gdp, w_gdp, tr_gdp, distw)) %>% 
  mutate(distw = as.numeric(distw))
```

#Exports-Imports
```{r}
tur_ven <- read_csv("tur_ven_62_18.csv")
tur_ven2018 <- tur_ven %>% 
  filter(year == 2018)
tur_ven_sections_2018 <- tur_ven2018 %>% 
  select(c("section_shortname_english", 
           "export_value_usd", "import_value_usd")) %>% 
  rename(section = section_shortname_english ) %>% 
  group_by(section) %>% 
  summarise(exports = sum(export_value_usd, na.rm = T),
            imports = sum(import_value_usd, na.rm = T)) %>% 
  filter(exports > 0, 
         imports > 0)
plot_turkey_export_2018_bar <- tur_ven_sections_2018 %>% 
  mutate(exports = exports / 1000000) %>% 
  ggplot(aes(reorder(section, -exports), exports))+
  geom_col(aes(fill = section)) + 
  labs(title = "Turkey's exports to Venezuela in 2018", 
       x = "", y = "", subtitle = "In million USD", caption = "Source: World Bank") + 
  theme(legend.position = "none",
    axis.text.x=element_text(angle=45, hjust=1))
plot_turkey_import_2018_bar <- tur_ven_sections_2018 %>% 
  mutate(imports = imports / 1000000) %>% 
  ggplot(aes(reorder(section, -imports), imports))+
  geom_col(aes(fill = section)) + 
  labs(title = "Turkey's imports to Venezuela in 2018", 
       x = "", y = "", subtitle = "In million USD", caption = "Source: World Bank") + 
  theme(legend.position = "none",
    axis.text.x=element_text(angle=45, hjust=1))
```

```{r}
plot_turkey_export_2018_bar
plot_turkey_import_2018_bar
```


# Gravity
## Creating FTA dummy variable:
```{r}
# eu countries iso 3 codes for FDA dummy
eu <- c("AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", 
        "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX",
        "MLT", "NLD", "POL", "PRT", "ROU", "SVK", "SVN", "ESP", "SWE", "GBR") %>%
  tolower()
```

```{r}
df5 <- df4 %>% 
  filter(year > 1961) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1995,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2007,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1997,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2015,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2010,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "tun", 1, 0), 0),
    FTA = ifelse(year > 2017,ifelse(code == "ven", 1, 0), 0)
  )
dist_n = 1.25
gr <- df5 %>%
  mutate(gravity = (1/w_gdp) * (tr_gdp*gdp) / distw^dist_n)
gravity_plot <- gr %>%
  mutate(total_trade =total_trade/ 1000000) %>%
  filter(year > 1961) %>%
  ggplot(aes(gravity, total_trade)) + 
  geom_point(aes(color = total_trade, size = gravity, alpha = distw)) + 
  geom_smooth(method = "lm", se = F, color = "red") + 
  theme_classic()
gravity_plot2 <- gr %>%
  mutate(total_trade =total_trade/ 1000000,
         dist_category = case_when(
           distw < 5000 ~ "Short Distance",
           distw < 10000 ~ "Middle Distance",
           distw > 10000 ~ "Long Distance"
         )) %>%
  filter(!is.na(dist_category)) %>%
  ggplot(aes(gravity, total_trade)) + 
  geom_point(aes(color = total_trade, size = gravity, alpha = distw)) + 
  geom_smooth(method = "lm", se = F, color = "red") + 
  facet_grid(~ dist_category, scales = "free_y")
gravity_ven <- gr %>%
  mutate(total_trade =total_trade/ 1000000) %>%
  filter(year == 2018,
         gravity < .1) %>%
  ggplot(aes(gravity, total_trade)) + 
  geom_point(aes(color = distw, size = gravity)) + 
  annotate("text", x = 0.05839045, y = 532118521/1000000, 
           label = "Venezuela", color = "brown") + 
  geom_abline(x = -61.23,slope = 1675.76, color = "pink", size = 2) +
  theme_classic() 
gr_ols <- gr %>%
  mutate(total_trade =total_trade/ 1000000) %>%
  filter(year > 1961) %>%
  lm(data = .,
     formula = total_trade ~ gravity) %>% 
  summary()
```

```{r}
df5
```

```{r}
gravity_plot
gravity_plot2
gravity_ven
```




### After this part, the econometric models depends on the changes on this datasets.
### So things will start to be messy.

```{r}

```

# Econometrical Analysis

```{r}
## 62 
# ven not included
df5 <- df4 %>% 
  filter(year > 1961) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1995,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2007,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1997,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2015,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2010,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "tun", 1, 0), 0)
  )

df6 <- pdata.frame(df5, index = c("code", "year"))

No_Ven_62 <- plm(log(total_trade) ~ log(gdp) + log(tr_gdp) + log(distw) + FTA,
                 data = df6,model = "random")
# ven included

df5 <- df4 %>% 
  filter(year > 1961) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1995,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2007,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1997,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2015,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2010,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "tun", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "ven", 1, 0), 0)
  )

df6 <- pdata.frame(df5, index = c("code", "year"))

Yes_Ven_62 <- plm(log(total_trade) ~ log(gdp) + log(tr_gdp) + log(distw) + FTA,
                 data = df6,model = "random")

## 83
df5 <- df4 %>% 
  filter(year > 1983) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1995,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2007,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1997,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2015,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2010,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "tun", 1, 0), 0)
  )

df6 <- pdata.frame(df5, index = c("code", "year"))

No_Ven_80 <- plm(log(total_trade) ~ log(gdp) + log(tr_gdp) + log(distw) + FTA,
                 data = df6,model = "random")
# ven included

df5 <- df4 %>% 
  filter(year > 1983) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1995,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2007,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1997,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2015,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2010,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "tun", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "ven", 1, 0), 0)
  )

df6 <- pdata.frame(df5, index = c("code", "year"))

Yes_Ven_80 <- plm(log(total_trade) ~ log(gdp) + log(tr_gdp) + log(distw) + FTA,
                 data = df6,model = "random")

## 02
df5 <- df4 %>% 
  filter(year > 2002) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1995,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2007,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1997,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2015,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2010,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "tun", 1, 0), 0)
  )

df6 <- pdata.frame(df5, index = c("code", "year"))

No_Ven_00 <- plm(log(total_trade) ~ log(gdp) + log(tr_gdp) + log(distw) + FTA,
                 data = df6,model = "random")
# ven included

df5 <- df4 %>% 
  filter(year > 2002) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1995,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2007,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1997,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2015,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2010,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "tun", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "ven", 1, 0), 0)
  )

df6 <- pdata.frame(df5, index = c("code", "year"))

Yes_Ven_00 <- plm(log(total_trade) ~ log(gdp) + log(tr_gdp) + log(distw) + FTA,
                 data = df6,model = "random")

stargazer(No_Ven_62, Yes_Ven_62, No_Ven_80, Yes_Ven_80, No_Ven_00, Yes_Ven_00,
          type = "text", 
          column.labels = c("No Ven 62","Yes Ven 62","No Ven 80","Yes Ven 80","No Ven 00","Yes Ven 00"),
          object.names = T)
```

#the answers to the last two questions
## adding VEN_TUR as dummy 
```{r}

df5 <- df4 %>% 
  filter(year > 1962) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1995,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2007,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1997,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2015,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2016,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2010,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "tun", 1, 0), 0),
    VEN_TUR = ifelse(year > 2016,ifelse(code == "ven", 1, 0), 0)
  )

df6 <- pdata.frame(df5, index = c("code", "year"))

Add_Ven_Tur<- plm(log(total_trade) ~ log(gdp) + log(tr_gdp) + log(distw) + FTA + VEN_TUR,
                 data = df6,model = "random")


stargazer(Add_Ven_Tur, type = "text")
```

## making time lagged (2 years)
```{r}
df5 <- df4 %>% 
  filter(year > 1962) %>%
  mutate(
    FTA = 0,
    FTA = ifelse(year > 1993,ifelse(code %in% eu, 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "alb", 1, 0), 0),
    FTA = ifelse(year > 2001,ifelse(code == "bih", 1, 0), 0),
    FTA = ifelse(year > 2005,ifelse(code == "egy", 1, 0), 0),
    FTA = ifelse(year > 2009,ifelse(code == "chl", 1, 0), 0),
    FTA = ifelse(year > 2006,ifelse(code == "geo", 1, 0), 0),
    FTA = ifelse(year > 1995,ifelse(code == "isr", 1, 0), 0),
    FTA = ifelse(year > 2013,ifelse(code == "mys", 1, 0), 0),
    FTA = ifelse(year > 2014,ifelse(code == "mda", 1, 0), 0),
    FTA = ifelse(year > 2004,ifelse(code == "mar", 1, 0), 0),
    FTA = ifelse(year > 2008,ifelse(code == "srb", 1, 0), 0),
    FTA = ifelse(year > 2011,ifelse(code == "kor", 1, 0), 0),
    FTA = ifelse(year > 2003,ifelse(code == "tun", 1, 0), 0),
    VEN_TUR = ifelse(year > 2016,ifelse(code == "ven", 1, 0), 0)
  )

df6 <- pdata.frame(df5, index = c("code", "year"))

Add_Ven_Tur<- plm(log(total_trade) ~ log(gdp) + log(tr_gdp) + log(distw) + FTA + VEN_TUR,
                 data = df6,model = "random")

stargazer(Add_Ven_Tur, type = "text")
```


#Labor Analysis
```{r}
DvsS <- WDI(country = c("TUR", "VEN"),
            indicator = c("SL.TLF.TOTL.IN",
                          "SL.TLF.TOTL.FE.ZS"))
DS <- DvsS %>%
  filter(year > 1990) %>%
  mutate(
    LaborForce = SL.TLF.TOTL.IN / 1000000 ,
    LPer_FE = round(SL.TLF.TOTL.FE.ZS, digits = 2) / 100,
    Labor_FE = LaborForce * LPer_FE,
    Labor_MA = LaborForce - Labor_FE) %>%
  select(-c(iso2c, SL.TLF.TOTL.IN, SL.TLF.TOTL.FE.ZS))
DvsS2 <- WDI(country = c("TUR", "VEN"),
            indicator = c("SL.AGR.EMPL.MA.ZS",
                          "SL.IND.EMPL.MA.ZS",
                          "SL.SRV.EMPL.MA.ZS",
                          "SL.AGR.EMPL.FE.ZS",
                          "SL.IND.EMPL.FE.ZS",
                          "SL.SRV.EMPL.FE.ZS"))
DS2 <- DvsS2 %>% filter(year > 1990) %>%
  mutate(
    AGRPer_MA = round(SL.AGR.EMPL.MA.ZS, digits = 2) / 100,
    INDPer_MA = round(SL.IND.EMPL.MA.ZS, digits = 2) / 100,
    SRVPer_MA = round(SL.SRV.EMPL.MA.ZS, digits = 2) / 100,
    AGRPer_FE = round(SL.AGR.EMPL.FE.ZS, digits = 2) / 100,
    INDPer_FE = round(SL.IND.EMPL.FE.ZS, digits = 2) / 100,
    SRVPer_FE = round(SL.SRV.EMPL.FE.ZS, digits = 2) / 100
  ) %>%
  select(-c(iso2c, SL.AGR.EMPL.MA.ZS, SL.IND.EMPL.MA.ZS,SL.SRV.EMPL.MA.ZS,
            SL.AGR.EMPL.FE.ZS, SL.IND.EMPL.FE.ZS,SL.SRV.EMPL.FE.ZS,
            country, year))

DS3 <- cbind(DS, DS2)
DS4 <- DS3 %>%
  mutate(
    AGR_MA = Labor_MA * AGRPer_MA,
    IND_MA = Labor_MA * INDPer_MA,
    SRV_MA = Labor_MA * SRVPer_MA,
    AGR_FE = Labor_FE * AGRPer_FE,
    IND_FE = Labor_FE * INDPer_FE,
    SRV_FE = Labor_FE * SRVPer_FE) %>% 
  reshape::melt(id = c("country", "year")) %>% 
  separate(
  col = variable,
  into = c("ind", "Gender"),
  sep = "_") %>%
  mutate(
    ind = as.factor(ind),
    Gender = as.factor(Gender))
  
plot_labor <- DS4 %>%
  filter(ind == "LaborForce") %>%
  ggplot()+
  geom_line(aes(x = year, y = value, color = country), 
           stat = "identity",position = "dodge")

```

```{r}
plot_labor
```

